---
title: "taylorDiD: modern DiD and event-study helpers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{taylorDiD walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 4.5
)
```

`taylorDiD` consolidates a set of difference-in-differences and event-study
helpers behind one consistent interface. It wraps three estimators:

- **BJS imputation** (Borusyak, Jaravel, and Spiess) via `didimputation`,
- **dCDH** (de Chaisemartin and D'Haultfoeuille) via `DIDmultiplegtDYN`, and
- **did2s** (Gardner two-stage) via `did2s`.

Every estimator returns the same result object and shares one set of house-style
conventions: unit-clustered standard errors, a joint pre-trends F-test, a
post-treatment average effect, the pre-treatment outcome mean, and matching plot
and LaTeX-table builders.

```{r}
library(taylorDiD)

panel <- readRDS(system.file("extdata", "synthetic_panel.rds",
                             package = "taylorDiD"))
head(panel)
```

## The example panel

One row is a unit-year. `id` is the unit, `year` the period, `g` the treatment
cohort (the year a unit first adopts; `Inf` for never-treated), `treat` the
binary absorbing treatment, `d` a non-binary treatment intensity, and `y` the
outcome. The data are simulated with a positive post-treatment effect that ramps
with event time.

```{r}
str(panel)
table(cohort = panel$g[!duplicated(panel$id)])
```

## BJS imputation event study

`did.event.study()` returns coefficient-by-event-time estimates, the joint
pre-trends F-test, and the post-treatment average effect (computed via an
auxiliary static DiD on the never-/not-yet-/treated-in-window subset).

```{r}
bjs <- did.event.study(panel, yname = "y", gname = "g", tname = "year",
                       idname = "id", pre.window = -4:-1, post.window = 0:3,
                       verbose = FALSE)
bjs$coefficients[, .(event.time, estimate, std.error, p.value)]
bjs$avg.post.effect$estimate
bjs$pretrends.test$p.value
```

Plot it with the house defaults (point estimates, 95% CI errorbars, dashed zero
line, serif theme):

```{r}
plot_event_study(bjs, y.title = "Effect on y")
```

The two-panel LaTeX table is produced by `tex_event_study_table()` (Panel A =
coefficients by event time; Panel B = average effect, pre-trends p-value,
pre-treatment mean, N):

```{r}
cat(tex_event_study_table(bjs, model.names = "BJS"), sep = "\n")
```

## dCDH event study

`did.event.study.dyn()` handles binary, discrete, and continuous treatments. It
takes the number of pre/post periods directly. (DIDmultiplegtDYN needs the
`polars` package; the wrapper attaches it for you.)

```{r, eval = requireNamespace("polars", quietly = TRUE), results = "hide"}
dcdh <- did.event.study.dyn(panel, yname = "y", dname = "treat", tname = "year",
                            idname = "id", n.pre = 3, n.post = 3,
                            verbose = FALSE)
```

```{r, eval = requireNamespace("polars", quietly = TRUE)}
dcdh$coefficients[, .(event.time, estimate, std.error, p.value)]
plot_event_study(dcdh, y.title = "Effect on y")
```

For a non-binary treatment, pass the intensity variable and `normalized = TRUE`
to interpret the coefficient as the effect of a one-unit increase:

```{r, eval = requireNamespace("polars", quietly = TRUE), results = "hide"}
dcdh_cont <- did.event.study.dyn(panel, yname = "y", dname = "d", tname = "year",
                                 idname = "id", n.pre = 3, n.post = 3,
                                 normalized = TRUE, verbose = FALSE)
```

> Note: the dCDH post-treatment average is the mean of the dynamic effects (the
> BJS and did2s estimators instead use an auxiliary static DiD). The dCDH
> estimator has no clean static-DiD analog, so its natural aggregate is the mean
> of the period-specific effects.

## did2s (Gardner two-stage)

`did2s_event_study()` runs the full two-stage flow: prepare the panel, fit
`did2s::did2s()` with unit and time fixed effects in the first stage and
event-time dummies in the second, then summarize with the same joint pre-trends
F-test and auxiliary-static-DiD post-average as the BJS estimator. Never-treated
units use the shared coding (`NA`/`Inf`/non-positive). The omitted reference is
event time -1.

```{r}
es_did2s <- did2s_event_study(panel, yname = "y", gname = "g", tname = "year",
                              idname = "id", pre.window = -4:-1,
                              post.window = 0:3, verbose = FALSE)
es_did2s$coefficients[, .(event.time, estimate, std.error, p.value)]
```

```{r}
plot_event_study(es_did2s, y.title = "Effect on y")
```

```{r}
cat(tex_event_study_table(es_did2s, model.names = "did2s"), sep = "\n")
```

## Shared conventions

- **Never-treated coding.** `is_never_treated()` treats `NA`, `Inf`, and
  non-positive timing as never-treated; `code_never_treated()` normalizes to a
  chosen sentinel. This is the single source of truth across estimators.
- **Joint pre-trends F-test.** `joint_pretrends_F()` reports Wald / n_pre with a
  p-value from an F with `df2 = Inf`. It uses the full covariance when available
  (did2s) and a diagonal approximation otherwise (BJS/dCDH).
- **Result object.** Every event study returns a `taylorDiD_es` object with
  `coefficients`, `pretrends.test`, `avg.post.effect`, `pre.mean.treated`, and
  the sample sizes -- consumed directly by the plot and table builders.

```{r}
class(bjs)
names(bjs)
```
